AiPoker - Implementation of a Poker Agent

### 2014-11-25 report/paper.tex
push

\usepackage{float}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
\usepackage{balance}
\usepackage[nooneline,hang,it,IT]{subfigure}

@@

\begin{document}
\pagenumbering{roman}
%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

@@

One possible use for Monte Carlo methods is to determine the value of an integral as illustrated in fig(\ref{fig:integralex}). By randomly placing dots and label them as either under or over the function value an estimate of the area can be achieved. By taking the area of the square encapsulating the function times the ratio between the number of points under the function and the total amount of points. The approximated value given by a Monte Carlo method will converge to the correct value as the number of sample points goes to infinity. [REF!]

@@

\textbf{Pre Flop}\\*
In this stage the players get two cards each from the dealer. These cards are called the pocket cards and after the cards have been dealt, the first betting round begins. If a player bets or raises all other players get the opportunity to either call, raise or fold max three times, after that the players who have not folded move on to the flop.
\\*\\*
\textbf{Flop}\\*
The flop is when the dealer deals the three first cards on the table. These cards are visible to all players. The pocket cards can now be combined with the flop cards in order to create a full hand. The players will make moves in the same procedure as in the pre flop stage.
\\*\\*
\textbf{Turn}\\*
Now the dealer adds a fourth card to the table. The players will now bet again.
\\*\\*
\textbf{River}\\*
The dealer adds a fifth card to the table. This is the last card which is dealt. After the last betting round, the game moves on to showdown. 
\\*\\*
\textbf{Showdown}\\*
In this phase the final betting is completed and the players who have not folded yet will now show their cards in order to decide who is the winner of this round. The winner will get the full pot and then a new round will start with the pre flop.
\\*\\*
\textbf{The different five combination of cards}\\*
There are nine different types of combination of cards and here they are listed in descending order:
\begin{itemize}
  \item \textit{Straight Flush}: When the player has five cards in a consecutive order of value within the same suit. If more than one player have a Straight Flush the player with the highest values of his/her cards wins. The best variant is the Royal Straight Flush when the player has Ace, King, Queen, Jack and a ten within the same suit.
  \item \textit{Four of a kind}: It is a hand that contains four of the same value in all the different suits. If more players have the same type of hand the hand with the highest four of a kind wins.
  \item \textit{Full House}: This hand is a combination of three of a kind and a pair. If more than one player have a Full House the player with the highest three of a kind wins. If more players have the same three of a kind, the player with the highest pair wins.
  \item \textit{Flush}: A flush is when a player has five cards within the same suit, if more than one player have a flush the one with the highest card wins. If they have the same highest card then the the second highest card is checked and so on.
  \item \textit{Straight}: When you have five cards in a consecutive order of value in different suits. If more than one player has a Straight then the player with the highest card wins.
  \item \textit{Three of a Kind}: When you have three cards of the same value, if more than one player has a Three of a kind then the player with the highest value of the cards wins.
  \item \textit{Two Pair}: When a player has two pair of cards, if more than one player have a two pair the player with the highest pair has the best hand between them.
  \item \textit{One Pair}: When you have one pair of cards, if more than one player have a one pair the player with the highest pair has the best hand between them.
  \item \textit{High Card}: If the card does not match any of the types above then the card with the highest value is considered. 
\end{itemize}

@@

The above example shows a part of a tic-tac-toe game tree. Every node represents a possible state in the game with the root node being the current state and every edge is the action leading to that state. 
The size for a game tree grows very fast when the possible actions that can be taken in each state increase. For example the game tree for an empty tic-tac-toe board has $255168$ leaf nodes. Two player Texas Hold'em has a search space size of $O(10^{18})$. [REF!]

@@

Expectiminimax is the continuation of the Minimax algorithm that allows for stochastic events in the game tree such as drawing cards in a game of poker. Expectiminimax trees add an additional type of nodes, chance nodes. At chance nodes a child $c$ is chosen with some probability $P(c)$. This means that a chance node for drawing a card from a deck of cards has $52$ children with a uniform probability distribution of $1/52$. The value for a chance node is the weighted average of all its children. [REF!]

@@

One simple approach is to model each opponent as a straight forward player, that means that we assume that an aggressive move (raise, bet) indicates a strong hand, and a passive move (check, call) indicates a weak hand. This can be implemented by holding a histogram with all stages (pre-flop, flop, turn, river) and all the actions taken by the player in those stages (fold, bet, call, raise, check). A raise in a game state indicates that a player has a strong hand if he usually does not raise in that state and vice versa. If the game goes to showdown it is possible to calculate the hand strength of the opponents cards on hand and thus making it possible to associate a hand strength with the actions taken that round. [REF!]

>>>>

\usepackage{fancyhdr}
\usepackage{float}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
\usepackage{balance}
\usepackage[nooneline,hang,it,IT]{subfigure}
\usepackage{lastpage}

@@

\begin{document}
\pagestyle{fancy}
\rhead{TNM095}
\lhead{\empty}
\cfoot{\empty}
\rfoot{Page \ \thepage \ of \ \pageref{LastPage}}
\pagenumbering{arabic}
%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

@@

One possible use for Monte Carlo methods is to determine the value of an integral as illustrated in fig(\ref{fig:integralex}). By randomly placing dots and label them as either under or over the function value an estimate of the area can be achieved. By taking the area of the square encapsulating the function times the ratio between the number of points under the function and the total amount of points. The approximated value given by a Monte Carlo method will converge to the correct value as the number of sample points goes to infinity\cite{Kleij}.

@@

\textbf{Pre Flop}\\*
In this stage the players get two cards each from the dealer. These cards are called the pocket cards and after the cards have been dealt, the first betting round begins. If a player bets or raises all other players get the opportunity to either call, raise or fold max three times, after that the players who have not folded move on to the flop.
\\*\\*
\textbf{Flop}\\*
The flop is when the dealer deals the three first cards on the table. These cards are visible to all players. The pocket cards can now be combined with the flop cards in order to create a full hand. The players will make moves in the same procedure as in the pre-flop stage.
\\*\\*
\textbf{Turn}\\*
Now the dealer adds a fourth card to the table. The players will now bet again.
\\*\\*
\textbf{River}\\*
The dealer adds a fifth card to the table. This is the last card which is dealt. After the last betting round, the game moves on to showdown. 
\\*\\*
\textbf{Showdown}\\*
In this phase the final betting is completed and the players who have not folded yet will now show their cards in order to decide who is the winner of this round. The winner will get the full pot and then a new round will start with the pre flop.
\\*\\*
\textbf{The different five combination of cards}\\*
There are nine different types of combination of cards and here they are listed in descending order:
\begin{itemize}
  \item \textit{Straight Flush}: When the player has five cards in a consecutive order of value within the same suit. If more than one player have a Straight Flush the player with the highest values of his/her cards wins. The best variant is the Royal Straight Flush when the player has Ace, King, Queen, Jack and a ten within the same suit.
  \item \textit{Four of a kind}: It is a hand that contains four of the same value in all the different suits. If more players have the same type of hand the hand with the highest four of a kind wins.
  \item \textit{Full House}: This hand is a combination of three of a kind and a pair. If more than one player have a Full House the player with the highest three of a kind wins. If more players have the same three of a kind, the player with the highest pair wins.
  \item \textit{Flush}: A flush is when a player has five cards within the same suit, if more than one player have a flush the one with the highest card wins. If they have the same highest card then the the second highest card is checked and so on.
  \item \textit{Straight}: When you have five cards in a consecutive order of value in different suits. If more than one player has a Straight then the player with the highest card wins.
  \item \textit{Three of a Kind}: When you have three cards of the same value, if more than one player has a Three of a kind then the player with the highest value of the cards wins.
  \item \textit{Two Pair}: When a player has two pair of cards, if more than one player have a two pair the player with the highest pair has the best hand between them.
  \item \textit{One Pair}: When you have one pair of cards, if more than one player have a one pair the player with the highest pair has the best hand between them.
  \item \textit{High Card}: If the card does not match any of the types above then the card with the highest value is considered. 
\end{itemize}

@@

The above example shows a part of a tic-tac-toe game tree. Every node represents a possible state in the game with the root node being the current state and every edge is the action leading to that state. 
The size for a game tree grows very fast when the possible actions that can be taken in each state increase. For example the game tree for an empty tic-tac-toe board has $255168$ leaf nodes. Two player Texas Hold'em has a search space size of $O(10^{18})$\cite{billingsdavidson}.

@@

Expectiminimax is the continuation of the Minimax algorithm that allows for stochastic events in the game tree such as drawing cards in a game of poker. Expectiminimax trees add an additional type of nodes, chance nodes. At chance nodes a child $c$ is chosen with some probability $P(c)$. This means that a chance node for drawing a card from a deck of cards has $52$ children with a uniform probability distribution of $1/52$. The value for a chance node is the weighted average of all its children\cite{Kleij}.

@@

One simple approach is to model each opponent as a straight forward player, that means that we assume that an aggressive move (raise, bet) indicates a strong hand, and a passive move (check, call) indicates a weak hand. This can be implemented by holding a histogram with all stages (pre-flop, flop, turn, river) and all the actions taken by the player in those stages (fold, bet, call, raise, check). A raise in a game state indicates that a player has a strong hand if he usually does not raise in that state and vice versa. If the game goes to showdown it is possible to calculate the hand strength of the opponents cards on hand and thus making it possible to associate a hand strength with the actions taken that round\cite{schweizer}.
